---
title: "model_diagnotics"
author: "Study group 3"
date: "`r Sys.Date()`"
output: html_document
---

# Set-up

```{r}
pacman::p_load(here, cmdstanr, tidyverse, posterior, bayesplot, rethinking)

library(cmdstanr)
```

```{r}
# import simulation data
data_raw <- read_csv("stan_test_data.csv")

# import model fit
fit <- readRDS("fitted_models/2ndassigment_fitted.rds")
```

# Basic model diagnostics

We will do the following checks: 

Visual
- Chain mixing (trace plots)

Numeric
- Divergent Transitions
- E-BFMI (Energy-Bayesian Fraction of Missing Information); below 0.3 is bad news
- Convergence checks (Rhats < 1.01)
- Effective Sample Size checks (ESS > 400, both bulk and tail)

```{r}
# trace plot of alpha and alpha_logit

# getting draws from the model
draws <- fit$draws(variables = c("alpha_logit", "alpha"))
bayesplot::mcmc_trace(draws) +
  labs(title = "Traceplots for MCMC sampling", x = "Sample", y = expression(alpha))
```
It looks nicely random!

Numeric checks:
To make sure everything is ok, we can use Riccardo's diagnostic battery from the course notes:


```{r}
# creating summaries
fit_summary <- fit$summary()
diagnostic_summary <- fit$diagnostic_summary()

# 1. Checking divergences
divergences <- sum(diagnostic_summary$num_divergent)
if (divergences > 0) {
  stop(paste("CRITICAL FAILURE:", divergences, "divergences detected. The posterior geometry is pathological. Reparameterization required. Do not interpret these results."))
}

# 2. Check E-BFMI (Energy-Bayesian Fraction of Missing Information)
# Threshold is typically 0.3. Values below this indicate poor exploration.
min_ebfmi <- min(diagnostic_summary$ebfmi)
if (min_ebfmi < 0.3) {
  warning(paste("WARNING: Low E-BFMI detected (Min =", round(min_ebfmi, 3), 
                "). The sampler may struggle to explore the tails of the posterior."))
}

# 3. Check for Convergence (Rhat < 1.01)
max_rhat <- max(fit_summary$rhat, na.rm = TRUE)
if (max_rhat > 1.01) {
  stop(paste("CRITICAL FAILURE: Chains did not converge. Max Rhat =", round(max_rhat, 3)))
}

# 4. Check Effective Sample Size (ESS > 400 for both bulk and tail)
min_ess_bulk <- min(fit_summary$ess_bulk, na.rm = TRUE)
min_ess_tail <- min(fit_summary$ess_tail, na.rm = TRUE)
if (min_ess_bulk < 400 || min_ess_tail < 400) {
  warning("WARNING: Insufficient Effective Sample Size. Autocorrelation is too high. Run longer chains.")
}
```

We can also plot the different diagnostics for an intuitive overview:

```{r}
print(diagnostic_summary)
hist(diagnostic_summary$ebfmi, xlim = c(0.3,max(diagnostic_summary$ebfmi)))
hist(fit_summary$rhat, xlim = c(min(fit_summary$rhat, na.rm = TRUE),1.01))
hist(fit_summary$ess_bulk, xlim = c(400, max(fit_summary$ess_bulk, na.rm = TRUE)+100))
```
No indication of problems with this initial fit!

We should probably do this for a 1000 fits though.. But maybe that is a part of later steps?


