---
title: "prior-posterior-update-check"
author: "Christian Stenbro"
date: "`r Sys.Date()`"
output: html_document
---

# set-up

```{r}
pacman::p_load(here, cmdstanr, tidyverse, posterior, bayesplot, rethinking)
```

```{r}
# import simulation data
data_raw <- read_csv("stan_test_data.csv")

# import model fit
fit <- readRDS("fitted_models/2ndassigment_fitted.rds")
```
# reasoning about prior-posterior update check

The goal is to investigate the relationship between the prior and the posterior. Concretely, we ask the following:

- Has the data been informative, meaning, *is the posterior distribution visually narrower than the prior distribution?*
- Is the prior too restrictive, meaning, *is the posterior distribution's mean different from the prior distribution's mean?*

The way we do this is by visualizing the marginal posterior distribution against the marginal prior distribution. The code is based on ch. 6 in Riccardo's book:

```{r}
# First, we extract the draws from the samples
draws <- as_draws_df(fit$draws())
```

```{r}
# Then, we define the ground truth for our estimand, alpha (the RL model learning rate. 
# This was set in our simulation to 0.2
true_alpha_prob <- 0.2
true_alpha_logit <- qlogis(true_alpha_prob) # notice that we are working in both the constrained cognitive space and the unlimited logit space simultaneously 
```

```{r}
# Unconstrained logit space
# Extracting the marginal prior and posterior distributions for alpha logit
update_logit_df <- tibble(
  parameter_value = c(draws$alpha_logit_prior, draws$alpha_logit),
  distribution = factor(
    rep(c("Prior", "Posterior"), each = nrow(draws)),
    levels = c("Prior", "Posterior") # these will be labels, needed for plotting
  )
)
```

```{r}
# lets look at the result
head(update_logit_df)
update_logit_df[10000:1010,] # just to check that we have posterior samples as well
```
```{r}
# plotting the marginal prior/posterior distribution in the unconstrained space
p_update_logit <- ggplot(update_logit_df, aes(x = parameter_value, fill = distribution)) +
  geom_density(alpha = 0.6, color = "white") +
  geom_vline(xintercept = true_alpha_logit, color = "black", linetype = "dashed", linewidth = 1) +
  scale_fill_manual(values = c("Prior" = "#ffbaba", "Posterior" = "skyblue")) +
  ggtitle("Prior-Posterior Update: Log-Odds Scale",
          subtitle = "Healthy = posterior (blue) narrower than prior (red), shifted toward true value (dashed)") +
  xlab(expression(alpha[logit] ~ " (Log-Odds)")) +
  ylab("Marginal Density") +
  annotate(
    "text", x = true_alpha_logit, y = Inf, 
    label = sprintf("Ground Truth = %s (round = 2)", round(true_alpha_logit, 2)), 
    vjust = 2, hjust = if (true_alpha_logit > 0) 1.1 else -0.1, 
    size = 4, fontface = "bold"
  ) +
  theme_classic() +
  theme(
    legend.position = "bottom", 
    legend.title = element_blank(),
    legend.text = element_text(size = 12)
  )

print(p_update_logit)
```
```{r}
# Making a similar plot in the constrained (probability) range

# Extracting the marginal prior and posterior distributions for alpha
update_prob_df <- tibble(
  parameter_value = c(draws$alpha_prior, draws$alpha),
  distribution = factor(
    rep(c("Prior", "Posterior"), each = nrow(draws)),
    levels = c("Prior", "Posterior") # these will be labels, needed for plotting
  )
)

# Plotting the marginal prior/posterior distribution in the constrained space
p_update_prob <- ggplot(update_prob_df, aes(x = parameter_value, fill = distribution)) +
  geom_density(alpha = 0.6, color = "white") +
  geom_vline(xintercept = true_alpha_prob, color = "black", linetype = "dashed", linewidth = 1) +
  scale_fill_manual(values = c("Prior" = "#ffbaba", "Posterior" = "skyblue")) +
  ggtitle("Prior-Posterior Update: Constrained Scale",
          subtitle = "Healthy = posterior (blue) narrower than prior (red), shifted toward true value (dashed)") +
  xlab(expression(alpha ~ "")) +
  ylab("Marginal Density") +
  annotate(
    "text", x = true_alpha_prob, y = Inf, 
    label = sprintf("Ground Truth = %s", true_alpha_prob), 
    vjust = 2, hjust = -0.1, 
    size = 4, fontface = "bold"
  ) +
  theme_classic() +
  theme(
    legend.position = "bottom", 
    legend.title = element_blank(),
    legend.text = element_text(size = 12)
  )

print(p_update_prob)

```

Plots are interpreted in the report. One note: The prior is cut of at the limits, but I think that is expected behaviour, exactly since we are in the constrained space. However, I am surprised by how artificial this cut-off appears.
