---
title: "SBC"
output: html_document
---

## Setup

```{r}
### YOU MIGHT NEED DEV VER OF SBC ### 
# devtools::install_github("hyunjimoon/SBC")

#load packages
pacman::p_load(tidyverse, cmdstanr, bayesplot, SBC, future)
```

## SBC Guide

You need a :
- To know what your mathematcial model is
- generator (r function) that simulates data according to the model
- and a **Backend**, a fitting procedure for fitting the mathematical model to data. (.stan file in this case)

SBC then checks when you backend (your model) and generator doesn't encode the same data generating process. First step is to setup the various parts:

Backend:
- cmdstan model
- setup SBC backend object

generator:
- load in opponent behavior
- agent function
- generator function
- turn into SBC object

```{r}

## BACKEND
#complile stan model
model <- cmdstan_model("2ndassignment_clean.stan")


#setup backend. Settings from course notes.
backend <- SBC_backend_cmdstan_sample(
  model = model,
  chains = 2,
  iter_warmup = 500,
  iter_sampling = 1000,
  refresh = 0, 
  show_messages = F
)


## Setup RL agent first
rl_agent_no_soft_max <- function(v_prev, choice_prev, outcome_prev, lr) {

  ## Get the right PE ## 
  #if last choice was right just the outcome as feedback
  if (choice_prev == 1){
    feedback = outcome_prev
  } 
  
  #otherwise flip outcome, so its used 
  else {
    feedback = 1 - outcome_prev
  }
  
  #get pe
  PE = feedback - v_prev
  
  #update expected value of playing right
  v_curr <- v_prev + lr*PE
  
  
  choice <- rbinom(1, 1, v_curr)
  
  
  #return results as named values
  return(list(choice = choice, v_curr = v_curr))
}

## GENERATOR
#load in simulated opponent data
ran_trials <- read_csv("stan_test_data.csv")

#make generator function
gen_fun <- function(){
  #draw prior
  alpha_logit = rnorm(n=1, mean=0, sd=1.5)
  alpha <- plogis(alpha_logit)
  
  #output result number of trials
  n_trials = sample(c(10, 25, 50, 100), 1)
  
  #read in bias_agent choices
  mismatcher_choices <- ran_trials$mismatcher_choices
  
  expected_values = c(.5)
    
  #prep vector for matcher (RL agent)
  matcher_choices = c(1)
  
  #setup vec for RL agent outcomes
  outcomes_matcher = c(if_else(mismatcher_choices [1] == matcher_choices[1],  1, 0))
  
  #loop over trial number
  for (t in 2:n_trials){
      
        ## Biased agent. MISMATCHER
        mismatcher_choice <- mismatcher_choices[t]
        
        ## RL part: MATCHER
        #get vals for previous trial for RL
        outcome_prev = outcomes_matcher[t-1]
        choice_prev = matcher_choices[t-1]
        expected_value_prev = expected_values[t-1]
        
        #make choice for rl agent
        result <- rl_agent_no_soft_max(expected_value_prev, choice_prev, outcome_prev, lr=alpha)
        
        matcher_choice <- result$choice
        v_curr  <- result$v_curr
        
        #store choices and expected values
        matcher_choices <- c(matcher_choices, matcher_choice)
        expected_values <- c(expected_values, v_curr)
      
        #get outcome at current trial and store
        outcome_matcher <- if_else(matcher_choice == mismatcher_choice, 1, 0)
        outcomes_matcher <- c(outcomes_matcher, outcome_matcher)
      }
  
  #list of vars / should expected val go in here
  output_list <- list(
    variables = list(alpha_logit = alpha_logit, alpha = alpha),
    generated = list(n = n_trials, choice = matcher_choices, reward = outcomes_matcher, prior_alpha_m = 0, prior_alpha_sd = 1.5))
  
  return(output_list)
}

#store it as proper sbc object
sbc_generator <- SBC_generator_function(gen_fun)
```
Then run simulations using backend and generator. Since this is expensive note that simulations are parrelized using future package. This is wrapped in a conditional that checks if simulations already exists or the regen flag is set to TRUE.
```{r}
### config
n_sims = 400 # 100 = 1min    on M1 macbook air
regen_flag = F


#make file place to store sbc results
sbc_results_file <- file.path("fitted_models/sbc_results.rds")

#check if file exists. If not, run sims
sim_results_exists <- file.exists(sbc_results_file)

if (!sim_results_exists | regen_flag){
  
    print("running simulations")
  
  # run simulations in parallel
  plan(multisession(workers = 2))
  
  #simulate
  sbc_datasets <- generate_datasets(sbc_generator, n_sims = n_sims)
  sbc_results <- compute_SBC(sbc_datasets, backend, keep_fits = F)
  
  #return to sequentil
  plan(sequential)
  
  #save object
  saveRDS(sbc_results, sbc_results_file)
  
} else { #Otherwise just load in the saved results
  
  print("loading existing file")
sbc_results <- readRDS(sbc_results_file)
}
  
```

After running sims, we plot the results.

```{r}
plot_rank_hist(sbc_results, variables = "alpha") +
  theme_classic()
ggsave("figures/sbc_rank_hist.png")
```
We can also plot the emperical CDF against the theoretical uniform
```{r}
plot_ecdf(sbc_results, variables = "alpha") +
  theme_classic()
ggsave("figures/sbc_ecdf.png")
```


We can also do classic median posterior against true value plots using the SBC results
```{r}

#get the estimates for alpha
alpha_results <- sbc_results$stats %>% 
  filter(variable == "alpha")  %>% 
  mutate(ci_width = q95-q5)

#add the n_trials by getting the n value from the generated data set
alpha_results$n_trials <- sapply(alpha_results$sim_id, function(i)     
  sbc_datasets$generated[[i]]$n)

#Draw plot
ggplot(alpha_results, aes(x=simulated_value, y = median, color = ci_width)) +
  geom_point() +
  geom_errorbar(aes(ymin = q5, ymax=q95), alpha = .5) +
  geom_abline(linetype = "dashed") +

  #facet by n trials
  facet_wrap(~n_trials) +
  
  #make it pretty
  theme_classic() +
  theme(legend.position = "bottom") +                                  
  labs(color = "95% CI width", x = "Fitted Alpha", y = "True Alpha")  

ggsave("figures/param_recov_plot.png")

```



